Dark Control: The Default Mode Network as a Reinforcement Learning Agent

### 2018-08-09 paper/paper.tex
interm

\begin{abstract}
The default mode network (DMN) is believed to subserve the
baseline mental activity in humans.
%
Its highest energy consumption compared to other brain networks and
its intimate coupling with conscious awareness are both pointing to
an overarching function.
%
Many research streams speak in favor of an evolutionarily adaptive role in
envisioning experience to anticipate the future.
In the present work, we propose a \textit{process model}
that tries to explain \textit{how}
the DMN may implement
continuous evaluation and prediction of the environment to guide behavior.
%
% DMN function is recast in terms of
% control theory and reinforcement learning by basing it on Markov decision processes.
%
The main purpose of DMN activity, we argue,
may be to optimize action policies via value estimates
based on Markov Decision Processes through vicarious trial and error.
%
Our formal perspective on DMN function naturally accommodates as special cases
previous interpretations based on
(1) predictive coding,
(2) semantic associations, and
(3) a sentinel role.
%
Moreover, this process model for the neural optimization of complex behavior in the DMN
offers parsimonious explanations for
recent experimental findings in animals and humans.
\end{abstract}

@@

Psychologically,
the more the ongoing executed task is unknown and unpracticed,
the less stimulus-independent thoughts occur
\citep{filler1973daydreaming, teasdale1995stimulus, christoff2016mind}.
Conversely,
it has been empirically shown that, the more the world is easy to foresee,
the more human mental activity
becomes detached from the actual sensory environment
\citep{antrobus1966studies, pope1978regulation, mason2007, weiss2006}.
Without requiring explicit awareness,
these ``offline'' processes may contribute
to optimizing control of the organism.
We formalize
a \textit{policy matrix} to capture the space of possible actions that
the agent can perform
on the environment given the current state. A \textit{value function}
maps environmental objects and events (i.e., states) to expected rewards.
Switching between states reduces to a sequential processing model.
Informed by outcomes of performed actions, neural computation reflected in
DMN dynamics could be constantly shaped by prediction error through feedback loops.
The present computational account of DMN function will be described in the mathematical framework of Markov Decision Processes (MDP). MDPs specifically formalize decision making in stochastic contexts with reward feedback.

Such an RL account of DMN function can naturally embed human behavior
into the tension between exploitative action with immediate gains and
exploratory action with longer-term gratification.
We argue that DMN implication in many
of the most advanced human capacities
can be recast as prediction error minimization
informed by internally generated probabilistic simulations
- ``covert forms of action and perception'' \citep{pezzulo2011grounding} -,
allowing maximization of action outcomes across different time scales.
Such a purposeful optimization objective
may be solved by a stochastic approximation
based on a brain implementation of Monte Carlo sampling.
Even necessarily imperfect memory recall,
random day-time mind-wandering, and
seemingly arbitrary dreams during sleep
may provide randomly sampled blocks of pseudo-experience  that are
instrumental to iteratively
optimize the behavioral agenda of the organism.

@@

Please appreciate the importance of differentiating which levels of observation are at play in the present account. A process model is not solely intended to capture behavior of the agent, such as cognitive accounts of DMN function, but also the neurocomputational specifics of the agent. Henceforth, we will use ``inference`` when referring to aspects of the statistical model, ``prediction`` when referring to the neurobiological implementation, and words like ``forecast`` or ``forsee`` when referring to the behavior of the agent.
%
It is moreover important to note that
our account does not claim that neural activity in the DMN in particular
or the brain in general are identical with reinforcement learning algorithms.
Rather, we advocate feedback-based learning strategies as an attractive alternative perspective
to describe, quantify, and interpret DMN function in neuroscience research.

@@

Normal and disturbed metabolic fluctuations in the
human PMC have been closely related to
changes of conscious awareness \citep{cavanna2006precuneus}.
Indeed,
the PMC matures relatively late (i.e., myelination) during postnatal development in monkeys
\citep{goldman1987development}, which is generally considered to
be a sign of evolutionary sophistication.
%
This DMN region has long been speculated to
reflect constant computation of
environmental statistics and its internal representation
as an inner ``mind's eye'' \citep{cavanna2006precuneus, leech_pcc2014}.
For instance, B\'alint's syndrome is a neurological disorder of conscious
awareness that can result from tissue damage in the posterior medial cortex
\citep{balint1909seelenlahmung, randy2008}.
Such neurological patients are plagued by an
inability to bind various individual features of the visual
environment into an integrated whole (i.e., simultanagnosia)
as well as an inability to direct action towards
currently unattended environmental objects
(i.e., optic ataxia).
%
Scanning complex scenes is impaired in that statistic or moving objects in the
environment may be invisible or disappear in the subject perception of the
patient \citep{mesulam_book, blumenfeld_book}.
%
This dysfunction can be viewed as a high-level impairment in gathering
information about alternative objects (i.e., exploration) as well as
using these environmental opportunities towards a behavioral goal (i.e., exploitation).
Congruently,
the human PMC was coupled in two different functional connectivity analyses
\citep{bzdok2015subspecialization}
with the amygdala, involved in significance evaluation, and
the nucleus accumbens (NAc), involved in reward evaluation.
Specifically, among all parts of the PMC,
the ventral posterior cingulate cortex was
most connected to the laterobasal
nuclei group of the amygdala
\citep{bzdok2015subspecialization}.
This amygdalar subregion has been proposed to
continuously scan environmental input
for biological relevance assessment
\citep{amygdala_bzdok, ghods2009fundamental, baxter2002amygdala}.

@@

\subsection{The prefrontal cortex: action consideration and stimulus-value association}
Analogous to the PMC,
the dorsomedial PFC (dmPFC) of the DMN is believed to subserve
multi-sensory processes
across time, space, and content domains to
exert top-level control on behavior.
Comparing to the PMC, however,
dmPFC function may be closer to a
``mental sketchpad'' \citep{goldman1996prefrontal}, as this DMN part
potentially subserves the de-novo construction and manipulation
of meaning representations instructed by stored semantics and memories
\citep{bzdok2013segregation}.
The dmPFC may subserve representation and assessment
of one's own and other individuals' action considerations - a necessary
component of a full-blown RL agent.
Generally,
neurological patients with tissue damage in the prefrontal cortex
are known to struggle with
adaptation to new stimuli and events
\citep{stuss1986frontal}.
Specifically, neural activity in the human dmPFC
reflected expectations about other peoples' actions and
outcomes of these predictions.
Neural activity in the dmPFC indeed explained the performance decline
of inferring other peoples' thoughts in aging humans \citep{moran2012social}.
Certain dmPFC neurons in macaque monkeys exhibited a preference
for processing others', rather than own, action
with fine-grained adjustment of contextual aspects \citep{yoshida2010neural}.
% In fact, the spatially neighboring dorsal anterior cingulate cortex
% has also been linked to computing values and efforts of
% persisting a behavioral plan versus switching the
% environmental context in several lesion studies \citep{kolling2016value}.
%
% Such highly abstract neural computations necessarily rely on the
% construction of probabilistic internal information drawing from
% memory recall, mental scene elaboration processes,
% and stored knowledge of world regularities.
% %
% Moreover,
% in a computational brain-imaging experiment,
% dorsomedial PFC activity preferentially processed the consequences of
% action choices that were considered but not actually executed,
% whereas ventromedial PFC (vmPFC) activity
% processed especially value outcomes of actions actually performed on the environment
% \citep{nicolle2012agent}.

Comparing to the dmPFC,
the vmPFC is probably more specifically devoted to
subjective value evaluation and
risk estimation of relevant environmental stimuli
(Fig. \ref{fig:neurobiological_overview_DMN} and \ref{fig:dmn}).
The ventromedial prefrontal DMN may
subserve adaptive behavior by bottom-up-driven
processing of “what matters now”,
drawing on sophisticated value representations
\citep{rolls_OFC, doherty2015structure}.
Quantitative lesion findings across 344 human individuals confirmed
a substantial impairment in value-based action choice
\citep{glascher2012lesion}.
Indeed,
this DMN region is preferentially connected with reward-related and limbic regions.
The vmPFC is well known to have direct connections
with the NAc
in axonal tracing studies in monkeys \citep{haber1995orbital}.
Congruently, the gray-matter volume of the vmPFC and NAc
correlated with indices of value-guided behavior and reward attitudes
in humans
\citep{lebreton2009automatic}.
NAc activity is further thought to reflect reward prediction signals
from dopaminergic neurotransmitter pathways
\citep{schultz1998predictive}
that not only channel action towards basic survival needs but also
enable more abstract reward processings, and thus perhaps RL,
in humans \citep{doherty2015structure}.

@@

  \includegraphics[width=.75\linewidth]{fig_dmn.png}
  \caption{\textbf{Predictive structural association between reward system and
      DMN.} Reward tasks \citep{doherty2015structure} and neural processign in the DMN \citep{randy2008}, often call task-negative, have been studied so far in largely separate niches of the neuroscience literature. A currently underappreciated link is however suggested here based on 9,932 human subjects from the UK Biobank, inter-individual differences in left NAc volume ($R^2=0.11 \pm 0.02$ [standard deviation across cross-validation folds]) and right NAc volume ($R^2=0.14 \pm 0.02$) could be predicted from (z-scored) volume in the DMN regions. These out-of-sample generalizations reflect the expected performance in yet-to-be observed individuals obtained from linear support vector regression applied to normalized region volumes in the DMN in a 10-fold cross-validation procedure \citep{hastie}.
    Consistent for the left and right reward system, NAc volume in a given subject is
    positively coupled with the vmPFC and HC. The congruence of our structural association results for both NAc targets speaks to the robustness of our pattern-prediction findings. The opposite relation of the left and right TPJ to the NAc appears to reflect a repeatedly recognized hemispheric asymmetry with respect to functional implications \citep{seghier2013angular}, impairments in neurological patients \citep{corbetta2000}, different types of connectivity \citep{uddin_tpj, caspers2011} as well as micro- and macroanatomy \citep{caspers2006, caspers2008}. The colors are indicative of the (\textcolor{red}{red} = positive,
    \textcolor{blue}{blue} = negative) and relative importance (the lighter the higher) of the regression coefficients.
  The code for reproduction and visualization: \url{www.github.com/banilo/darkcontrol_2018}.}
  \label{fig:dmn}
% \end{wrapfigure}
\end{figure}

Consistently, diffusion MRI tractography in humans and monkeys
\citep{croxson2005quantitative}
quantified the NAc to
be more connected to the vmPFC than dmPFC in both species.
Two different functional connectivity analyses in humans also revealed
strong vmPFC connections
with the NAc, hippocampus (HC),
and PMC \citep{bzdok2015subspecialization}.
%
In line with these connectivity findings in animals and humans,
the vmPFC is often proposed to represent triggered
emotional and motivational states \citep{damasio1996somatic}.
Such real or imagined arousal states could be mapped in the vmPFC
as a bioregulatory disposition influencing cognition
and decision making.
In neuroeconomic studies of human decision making,
the vmPFC consistently reflects an individual’s subjective
value predictions
\citep{behrens2008associative},
which may also explain why performance within and across participants
was reported to relate to state encoding in the vmPFC \citep{Schuck2016}.
Such a ``cognitive map'' of the action space was argued to encode
the current task state even when states are unobservable from
the sensory environment.

@@

There is hence increasing evidence that
HC function extends beyond simple forms of
encoding and reconstruction of memory and space information.
Based on spike recordings of hippocampal neuronal populations,
complex spiking patterns can be followed across extended periods including
their modification of input-free self-generated patterns
after environmental events \citep{buzsaki2004large}.
Specific spiking sequences, which were elicited by experimental task design,
have been shown to be re-enacted spontaneously during
quiet wakefulness and sleep \citep{hartley2014space, o2010play}.
Moreover, neuronal spike sequences measured in hippocampal place cells of rats
featured re-occurrence directly after experimental trials
as well as directly before (prediction of) upcoming experimental trials \citep{diba2007forward}.
Similar spiking patterns in hippocampal neurons during rest and sleep
have been proposed to be critical in communicating local information
to the neocortex for long-term storage, potentially including DMN regions.
Moreover, in mice, invasively triggering spatial experience recall
in the HC during sleep
has been demonstrated to subsequently alter action choice during wakefulness
\citep{lavilleon2015}.
These HC-subserved mechanisms
conceivably contribute to advanced cognitive processes that require
re-experiencing or newly constructed mental scenarios,
such as in recalling autobiographical memory episodes
\citep{hassabis2007patients}.
Within a RL framework, the HC could thus orchestrate re-experience of environmental aspects for
consolidations based on re-enactment and for integration into
rich mental scene construction \citep{deuker2016event, bird2010establishing}.
As such, the HC may impact
ongoing perception of and action on the environment
\citep{maguire2016, lavilleon2015}.
% (TODO: relate to self-play in deepmind Go).
% (TODO: Kai had an interesting idea: the deja-vu effect
% could be when you already accidentally perturbed past
% such that it really occures in reality later.)

@@

The environment is generally taken as \textit{stochastic},
that is, changing in random ways. In addition, the environment is only
\textit{partially observable} in the sense that only limited aspects of the environment's
state are accessible to the agent's sensory input
\citep{starkweather2017dopamine}.
We assume that volatility of the environment
is realistic in a computational model which sets out
to explain DMN functions of the human brain.
% XXX: the following passage is already in the introduction
% Indeed, it is known that the more the external world is predictable,
% the more mental activity becomes detached from the actual sensory inputs
% \citep{antrobus1966studies, pope1978regulation},
% the more stimulus-independent thoughts occur, and
% the higher the neural activity in the DMN.
% Conversely, the more the currently executed task is unknown and unpracticed,
% the lower the DMN activity
% \citep{filler1973daydreaming, teasdale1995stimulus}.
% %
% These ``offline'' processes may however contribute to
% optimizing control of the organism.
% Such a framework of DMN funtion based on reinforcement learning
% can naturally embed human behavior
% in the tension between exploitative action with immediate gains and
% explorative action with longer-term reward schedules
% \citep{dayan2008decision}.
We argue that a functional account of the DMN based on RL
can naturally embed human behavior
in the tension between exploitative action with immediate gains and
explorative action with longer-term reward outcomes
\citep{dayan2008decision}.
In short, DMN implication in a diversity of
particularly sophisticated human behaviors
can be parsimoniously explained as instantiating probabilistic
simulations of experience coupled with prediction error minimization
to calibrate action trajectories for
reward outcome maximization at different time-scales.
Such a purposeful optimization objective
may be subserved by a stochastic approximation
based on a brain implementation of Monte Carlo sampling.

@@

\subsubsection{Does the hippocampus subserve Monte Carlo sampling?}
In RL, Monte Carlo simulation is a common means to update the agent's belief state
based on stochastic sampling around states and possible transitions \citep{daw2014algorithmic,silver2010monte}.
Monte Carlo simulation provides a simple method for evaluating the value of a state.
This inference procedure provides an effective mechanism both for tree search (of the considered
action trajectories)
and for belief state updates, breaking the curse of dimensionality and allowing much greater scalability than an RL agent without stochastic resampling procedures.
Such methods have scaling as a function of available data (i.e., sample complexity) that
is determined only by the underlying difficulty of the MDP, rather than the size of the state space or observation space,
which can be prohibitively large.

@@

\begin{figure}[!h]
  \includegraphics[width=.9\linewidth]{neurobiological_and_rl_overview_DMN.pdf}
  \caption{\textbf{Default mode network:
  possible neurobiological implementation of reinforcement learning.}
  Overview of how the constituent regions of the DMN (refer to section \ref{sec:nodes}; blue horizontal dashed line indicates the border between BA9 and BA10) may
  map onto computational components necessary for an RL agent.
  %
  Axonal tracing in monkeys and diffusion tractography in humans suggested
  that the NAc of the reward circuitry has monosynaptic fiber connections
  to the vmPFC \citep{haber1995, croxson2005quantitative}. Evaluation of
  propagated value information and triggered affective states encoded
  in the vmPFC may then feed into the functionally connected partner nodes of
  the DMN, such as the dmPFC and PMC \citep{andrews2010, bzdok2013segregation}.
}
  \label{fig:rl_process_chart}
\end{figure}

>>>>

\begin{abstract}
The default mode network (DMN) is believed to subserve the
baseline mental activity in humans.
%
Its higher energy consumption compared to other brain networks and
its intimate coupling with conscious awareness are both pointing to
an unknown overarching function.
%
Many research streams speak in favor of an evolutionarily adaptive role in
envisioning experience to anticipate the future.
In the present work, we propose a \textit{process model}
that tries to explain \textit{how}
the DMN may implement
continuous evaluation and prediction of the environment to guide behavior.
%
% DMN function is recast in terms of
% control theory and reinforcement learning by basing it on Markov decision processes.
%
The main purpose of DMN activity, we argue,
may be described by Markov Decision Processes
that optimize action policies via value estimates
based through vicarious trial and error.
%
Our formal perspective on DMN function naturally accommodates as special cases
previous interpretations based on
(1) predictive coding,
(2) semantic associations, and
(3) a sentinel role.
%
Moreover, this process model for the neural optimization of complex behavior in the DMN
offers parsimonious explanations for
recent experimental findings in animals and humans.
\end{abstract}

@@

At the psychological level,
the more the ongoing executed task is unknown and unpracticed,
the less stimulus-independent thoughts occur
\citep{filler1973daydreaming, teasdale1995stimulus, christoff2016mind}.
Conversely,
it has been empirically shown that, the more the world is easy to foresee,
the more human mental activity
becomes detached from the actual sensory environment
\citep{antrobus1966studies, pope1978regulation, mason2007, weiss2006}.
Without requiring explicit awareness,
these ``offline'' processes may contribute
to optimizing control of the organism in general.
We formalize
a \textit{policy matrix} to capture the space of possible actions that
the agent can perform
on the environment given the current state. A \textit{value function}
maps environmental objects and events (i.e., states) to expected reward outcomes.
Switching between states reduces to a sequential processing model.
Informed by outcomes of performed actions, neural computation reflected in
DMN dynamics could be constantly shaped by prediction error through feedback loops.
The present computational account of DMN function will be described in the mathematical framework of Markov Decision Processes (MDP). MDPs specifically formalize decision making in stochastic contexts with reward feedback.

Such a RL perspective on DMN function can naturally embed human behavior
into the tension between exploitative action with immediate gains and
exploratory action with longer-term gratification.
We argue that DMN implication in many
of the most advanced human capacities
can be recast as prediction error minimization
informed by internally generated probabilistic simulations
- ``covert forms of action and perception'' \citep{pezzulo2011grounding} -,
allowing maximization of action outcomes across different time scales.
Such a purposeful optimization objective
may be solved by a stochastic approximation
based on a brain implementation of Monte Carlo sampling.
Even necessarily imperfect memory recall,
random day-time mind-wandering, and
seemingly arbitrary dreams during sleep
may provide randomly sampled blocks of pseudo-experience  that are
instrumental to iteratively
optimize the behavioral agenda of the organism.

@@

Please appreciate the importance of differentiating which levels of observation are at play in the present account. A process model is not solely intended to capture behavior of the agent, such as cognitive accounts of DMN function, but also the neurocomputational specifics of the agent. Henceforth, we will use ``inference`` when referring to aspects of the statistical model, ``prediction`` when referring to the neurobiological implementation, and words like ``forecast`` or ``forsee`` when referring to the behavior of the agent.
%
It is moreover important to note that
our account does not claim that neural activity in the DMN in particular
or the brain in general are identical with reinforcement learning algorithms.
Rather, we advocate feedback-based learning strategies as an attractive alternative perspective
to describe, quantify, and interpret research findings on the DMN.

@@

Normal and disturbed metabolic fluctuations in the
human PMC have been closely related to
changes of conscious awareness \citep{cavanna2006precuneus, leech_pcc2014}.
Indeed,
the PMC matures relatively late (i.e., myelination) during postnatal development in monkeys
\citep{goldman1987development}, which is generally considered to
be a sign of evolutionary sophistication.
%
This DMN region has long been speculated to
reflect constant computation of
environmental statistics and its internal representation
as an inner ``mind's eye'' \citep{cavanna2006precuneus, leech_pcc2014}.
For instance, B\'alint's syndrome is a neurological disorder of conscious
awareness that can result from tissue damage in the posterior medial cortex
\citep{balint1909seelenlahmung, randy2008}.
Such neurological patients are plagued by an
inability to bind various individual features of the visual
environment into an integrated whole (i.e., simultanagnosia)
as well as an inability to direct action towards
currently unattended environmental objects
(i.e., optic ataxia).
%
Scanning complex scenes is impaired in that statistic or moving objects in the
environment may be invisible or disappear in the subject perception of the
patient \citep{mesulam_book, blumenfeld_book}.
%
This dysfunction can be viewed as a high-level impairment in gathering
information about alternative objects (i.e., exploration) as well as
using these environmental opportunities towards a behavioral goal (i.e., exploitation).
Congruently,
the human PMC was coupled in two different functional connectivity analyses
\citep{bzdok2015subspecialization}
with the amygdala, involved in significance evaluation, and
the nucleus accumbens (NAc), involved in reward evaluation.
Specifically, among all parts of the PMC,
the ventral posterior cingulate cortex was
most connected to the laterobasal
nuclei group of the amygdala
\citep{bzdok2015subspecialization}.
This amygdalar subregion has been proposed to
continuously scan environmental input
for biological relevance assessment
\citep{amygdala_bzdok, ghods2009fundamental, baxter2002amygdala}.

@@

\subsection{The prefrontal cortex: action consideration and stimulus-value association}
Analogous to the PMC,
the dorsomedial PFC (dmPFC, related to BA9) of the DMN is believed to subserve
multi-sensory processes
across time, space, and content domains to
exert top-level control on behavior.
Comparing to the PMC, however,
dmPFC function may be closer to a
``mental sketchpad'' \citep{goldman1996prefrontal}. This DMN part
potentially subserves the de-novo construction and manipulation
of meaning representations instructed by stored semantics and memories
\citep{bzdok2013segregation}.
The dmPFC may subserve representation and assessment
of one's own and other individuals' action considerations - a necessary
component of a full-blown RL agent.
Generally,
neurological patients with tissue damage in the prefrontal cortex
are known to struggle with
adaptation to new stimuli and events
\citep{stuss1986frontal}.
Specifically, neural activity in the human dmPFC
reflected expectations about other peoples' actions and
outcomes of these predictions.
Neural activity in the dmPFC indeed explained the performance decline
of inferring other peoples' thoughts in aging humans \citep{moran2012social}.
Certain dmPFC neurons in macaque monkeys exhibited a preference
for processing others', rather than own, action
with fine-grained adjustment of contextual aspects \citep{yoshida2010neural}.
% In fact, the spatially neighboring dorsal anterior cingulate cortex
% has also been linked to computing values and efforts of
% persisting a behavioral plan versus switching the
% environmental context in several lesion studies \citep{kolling2016value}.
%
% Such highly abstract neural computations necessarily rely on the
% construction of probabilistic internal information drawing from
% memory recall, mental scene elaboration processes,
% and stored knowledge of world regularities.
% %
% Moreover,
% in a computational brain-imaging experiment,
% dorsomedial PFC activity preferentially processed the consequences of
% action choices that were considered but not actually executed,
% whereas ventromedial PFC (vmPFC) activity
% processed especially value outcomes of actions actually performed on the environment
% \citep{nicolle2012agent}.

Comparing to the dmPFC,
the ventromedial PFC (vmPFC, related to BA10) is probably more specifically devoted to
subjective value evaluation and
risk estimation of relevant environmental stimuli
(Fig. \ref{fig:neurobiological_overview_DMN} and \ref{fig:dmn}).
The ventromedial prefrontal DMN may
subserve adaptive behavior by bottom-up-driven
processing of “what matters now”,
drawing on sophisticated value representations
\citep{rolls_OFC, doherty2015structure}.
Quantitative lesion findings across 344 human individuals confirmed
a substantial impairment in value-based action choice
\citep{glascher2012lesion}.
Indeed,
this DMN region is preferentially connected with reward-related and limbic regions.
The vmPFC is well known to have direct connections
with the NAc
in axonal tracing studies in monkeys \citep{haber1995orbital}.
Congruently, the gray-matter volume of the vmPFC and NAc
correlated with indices of value-guided behavior and reward attitudes
in humans
\citep{lebreton2009automatic}.
NAc activity is further thought to reflect reward prediction signals
from dopaminergic neurotransmitter pathways
\citep{schultz1998predictive}
that not only channel action towards basic survival needs but also
enable more abstract reward processings, and thus perhaps RL,
in humans \citep{doherty2015structure}.

@@

  \includegraphics[width=.75\linewidth]{fig_dmn.png}
  \caption{\textbf{Predictive structural association between reward system and
      DMN nodes.} Reward tasks \citep{doherty2015structure} and neural processing in the DMN \citep{randy2008}, often considered task-negative, have been studied so far in largely separate niches of the neuroscience literature. A currently underappreciated link is however suggested here based on 9,932 human subjects from the UK Biobank, inter-individual differences in left NAc volume ($R^2=0.11 \pm 0.02$ [standard deviation across cross-validation folds]) and right NAc volume ($R^2=0.14 \pm 0.02$) could be predicted from (z-scored) volume in the DMN regions. These out-of-sample generalizations reflect the expected performance in yet-to-be observed individuals obtained from linear support vector regression applied to normalized region volumes in the DMN in a 10-fold cross-validation procedure \citep{hastie}.
    Consistent for the left and right reward system, NAc volume in a given subject is
    positively coupled with the vmPFC and HC. The congruence of our structural association results for both NAc targets speaks to the robustness of our pattern-prediction findings. The opposite relation of the left and right TPJ to the NAc appears to reflect a repeatedly recognized hemispheric asymmetry with respect to functional implications \citep{seghier2013angular}, impairments in neurological patients \citep{corbetta2000}, different types of connectivity \citep{uddin_tpj, caspers2011} as well as micro- and macroanatomy \citep{caspers2006, caspers2008}. The colors are indicative of the (\textcolor{red}{red} = positive,
    \textcolor{blue}{blue} = negative) and relative importance (the lighter the higher) of the regression coefficients.
  The code for reproduction and visualization: \url{www.github.com/banilo/darkcontrol_2018}.}
  \label{fig:dmn}
% \end{wrapfigure}
\end{figure}

Consistently, diffusion MRI tractography in humans and monkeys
\citep{croxson2005quantitative}
quantified the NAc to
be more connected to the vmPFC than dmPFC in both species.
Two different functional connectivity analyses in humans also revealed
strong vmPFC connections
with the NAc, hippocampus (HC),
and PMC \citep{bzdok2015subspecialization}.
%
In line with these connectivity findings in animals and humans,
the vmPFC is often proposed to represent triggered
emotional and motivational states \citep{damasio1996somatic}.
Such real or imagined arousal states could be mapped in the vmPFC
as a bioregulatory disposition influencing cognition
and decision making.
In neuroeconomic studies of human decision making,
the vmPFC consistently reflects an individual’s subjective
value predictions
\citep{behrens2008associative}. This finding
may also explain why performance within and across participants
was reported to relate to state encoding in the vmPFC \citep{Schuck2016}.
Such a ``cognitive map'' of the action space - an integral part of a RL
agent -
was argued to encode
the current task state even when states are unobservable from
the sensory environment.

@@

There is hence increasing evidence that
HC function extends beyond simple forms of
encoding and reconstruction of memory and space information.
Based on spike recordings of hippocampal neuronal populations,
complex spiking patterns can be followed across extended periods including
their modification of input-free self-generated patterns
after environmental events \citep{buzsaki2004large}.
Specific spiking sequences, which were elicited by experimental task design,
have been shown to be re-enacted spontaneously during
quiet wakefulness and sleep \citep{hartley2014space, o2010play}.
Moreover, neuronal spike sequences measured in hippocampal place cells of rats
featured re-occurrence directly after experimental trials
as well as directly before (prediction of) upcoming experimental trials \citep{diba2007forward}.
Similar spiking patterns in hippocampal neurons during rest and sleep
have been proposed to be critical in communicating local information
to the neocortex for long-term storage, potentially including DMN regions.
Moreover, in mice, invasively triggering spatial experience recall
in the HC during sleep
has been demonstrated to subsequently alter action choice during wakefulness
\citep{lavilleon2015}.
These HC-subserved mechanisms
conceivably contribute to advanced cognitive processes that require
re-experiencing or newly constructed mental scenarios,
such as in recalling autobiographical memory episodes
\citep{hassabis2007patients}.
Within a RL framework, the HC could thus orchestrate re-experience of environmental aspects for
consolidations based on re-enactment and for integration into
rich mental scene construction \citep{deuker2016event, bird2010establishing}.
In this way, the HC may impact
ongoing perception of and action on the environment
\citep{maguire2016, lavilleon2015}.
% (TODO: relate to self-play in deepmind Go).
% (TODO: Kai had an interesting idea: the deja-vu effect
% could be when you already accidentally perturbed past
% such that it really occures in reality later.)

@@

The environment is generally taken as \textit{stochastic},
that is, changing in random ways. In addition, the environment is only
\textit{partially observable} in the sense that only limited aspects of the environment's
state are accessible to the agent's sensory input
\citep{starkweather2017dopamine}.
We assume that volatility of the environment
is realistic in a computational model which sets out
to explain DMN functions of the human brain.
% XXX: the following passage is already in the introduction
% Indeed, it is known that the more the external world is predictable,
% the more mental activity becomes detached from the actual sensory inputs
% \citep{antrobus1966studies, pope1978regulation},
% the more stimulus-independent thoughts occur, and
% the higher the neural activity in the DMN.
% Conversely, the more the currently executed task is unknown and unpracticed,
% the lower the DMN activity
% \citep{filler1973daydreaming, teasdale1995stimulus}.
% %
% These ``offline'' processes may however contribute to
% optimizing control of the organism.
% Such a framework of DMN funtion based on reinforcement learning
% can naturally embed human behavior
% in the tension between exploitative action with immediate gains and
% explorative action with longer-term reward schedules
% \citep{dayan2008decision}.
We argue that a functional account of the DMN based on RL
can naturally embed human behavior
in the tension between exploitative action with immediate gains and
explorative action with longer-term reward outcomes
\citep{dayan2008decision}.
In short, DMN implication in a diversity of
particularly sophisticated human behaviors
can be parsimoniously explained as instantiating probabilistic
simulations of experience coupled with prediction error minimization
to calibrate action trajectories for
reward outcome maximization at different time-scales.
Such a purposeful optimization objective
may be subserved by a stochastic approximation
based on a brain implementation for Monte Carlo sampling of events and outcomes.

@@

\subsubsection{Does the hippocampus subserve Monte Carlo sampling?}
In RL, Monte Carlo simulation is a common means to update the agent's belief state
based on stochastic sampling around states and possible transitions \citep{daw2014algorithmic,silver2010monte}.
Monte Carlo simulation provides a simple method for evaluating the value of a state.
This inference procedure provides an effective mechanism both for tree search (of the considered
action trajectories)
and for belief state updates, breaking the curse of dimensionality and allowing much greater scalability than a RL agent without stochastic resampling procedures.
Such methods have scaling as a function of available data (i.e., sample complexity) that
is determined only by the underlying difficulty of the MDP, rather than the size of the state space or observation space,
which can be prohibitively large.

@@

\begin{figure}[!h]
  \includegraphics[width=.9\linewidth]{neurobiological_and_rl_overview_DMN.pdf}
  \caption{\textbf{Default mode network:
  possible neurobiological implementation of reinforcement learning.}
  Overview of how the constituent regions of the DMN (refer to section \ref{sec:nodes}; blue horizontal dashed line indicates the border between BA9 and BA10) may
  map onto computational components necessary for a RL agent.
  %
  Axonal tracing in monkeys and diffusion tractography in humans suggested
  that the NAc of the reward circuitry has monosynaptic fiber connections
  to the vmPFC \citep{haber1995, croxson2005quantitative}. Evaluation of
  propagated value information and triggered affective states encoded
  in the vmPFC may then feed into the functionally connected partner nodes of
  the DMN, such as the dmPFC and PMC \citep{andrews2010, bzdok2013segregation}.
}
  \label{fig:rl_process_chart}
\end{figure}
